---
title: "Final Project - EDA"
author: "Colm Kennedy"
date: "2022-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(ggplot2)
library(moderndive)

all <- read.csv('all.csv')

head(all)

```

# Exploratory Data Analysis

### Population

The populations are skewed right, meaning there are fewer high populations. Most populations lie between 4.6 million and 32 million. Population does not seem correlated with life expectancy, r = 0.05.

```{r}
summary(all$population)

ggplot(data = all, mapping = aes(x = population)) +
  geom_histogram(color = 'white') +
  scale_x_continuous(labels = scales::comma)

ggplot(data = all, mapping = aes(x = population, y = lifeexp)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)



# Correlation
get_correlation(all, formula = lifeexp ~ population)
# very low corr life exp and pop
```


### Child Mortality ***

Child mortality and life expectancy have an extremely high negative correlation, r = -0.91. The plot illustrates a strong linear relationship. This is a very good indicator of life expectancy, and a great candidate for our model.

```{r}
summary(all$childmort)

ggplot(data = all, mapping = aes(x = childmort, y = lifeexp)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)

get_correlation(all, formula = lifeexp ~ childmort)
# Extremely strong correlation

```


### Income ***

Income and life expectancy are also highly correlated, r = 0.72. The relationship appears logarithmic, applying log to income appears to make the relationship linear. This is another good candidate for our model.

```{r}
# Income

summary(all$income)

ggplot(data = all, mapping = aes(x = income, y = lifeexp)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)


ggplot(data = all, mapping = aes(x = log(income), y = lifeexp)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)


get_correlation(all, formula = lifeexp ~ income)
# Strong correlation
```

### Income and Child Mortality

Income and child mortality appear to have a relationship with each other and may be interacting. They are negatively correlated, r = -0.64. After applying log to income, the relationship appears much more linear. It would be worth trying an interaction model with income and child mortality. 

```{r}
ggplot(data = all, mapping = aes(x = income, y = childmort)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)


ggplot(data = all, mapping = aes(x = log(income), y = childmort)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)

get_correlation(all, formula = childmort ~ income)

```

